---
title: "AlphaFold analysis"
author: "Hanhee Jo (PID# A59017994)"
format: html
toc: true
---

Here we demonstrate how to analyze and make sense of models from AlphaFold. We begin by reading all the model PDB files...

```{r}
library(bio3d)
```

PDB file names of my models
```{r}
files <- list.files("hiv_monomer_94b5b_1/", 
           pattern = ".pdb",
           full.names = T)
```


Align and superpose
```{r}
pdbs <- pdbaln(files, fit=TRUE, exefile="msa")
```

RMSD analysis

```{r}
rd <- rmsd(pdbs)
mean(rd)
```

```{r}
library(pheatmap)
pheatmap(rd)
```

```{r}
source( "https://tinyurl.com/newviewngl" )
library(NGLVieweR)
view.pdbs(pdbs)
```

# PCA

```{r}
pc <- pca(pdbs)
plot(pc)
```


# Redidue conservation from alignment file

AlphaFold writes out the MSA it calculated and used for structure prediction to A3M.

```{r}
aln_file <- list.files(path="hiv_monomer_94b5b_1/",
                       pattern=".a3m$",
                        full.names = TRUE)
aln_file
```

```{r}
aln <- read.fasta(aln_file[1], to.upper = TRUE)
```

```{r}
dim(aln$ali)
```

Score residue consevation:

```{r}
sim <- conserv(aln)
```

```{r}
con <- consensus(aln, cutoff = 0.9)
con$seq
```

Plot the conservation along the sequence/structure

```{r}
plot(sim, typ="h")
```

Let's look at these conserved positions in the structure:

```{r}
pdb <- read.pdb( files[1] )

view.pdb(pdb, backgroundColor = "pink",
         highlight = atom.select(pdb, resno=25:28),
         highlight.style = "spacefill" )
```

